### 2021-08-10 bias_correction.rmd
Population averaging seems to work!

---
title: "Bias correction in GLMs"
author: Bicko, Jonathan & Ben
date: "2021 Jun 21 (Mon)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")

library(glmmTMB)
library(dplyr)
library(ggplot2)
library(vareffects); varefftheme()
library(ggpubr)
set.seed(1011)

## Functions
plotEsize <- function(df, pos = 0.5, col_lab = ""){
	pos <- position_dodge(pos)
	p1 <- (ggplot(df, aes(x = reorder(term, -estimate), y = estimate, colour="Estimate"))
		+ geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size=0.2, position = pos)
		+ geom_hline(yintercept=0,lty=2, size=0.2)
		+ coord_flip()
		+ labs(x="", y = "Estimate")
	)
}
```

\section{Introduction}

Prediction of the variation in the response variable depends on whether the relationship between the response variable and the predictors is linear or nonlinear. For example, when response, $Y$, changes nonlinearly with the predictor variable, $X$, averaged response variable with respect to the predictor, $E(Y(X))$, does not necessarily equal the response at the mean of the predictor, $Y(E(X)$. This can be understood in relation to Jensen's inequality which states that, for a nonlinear function, $Y(X)$, then $E(Y(X)) > Y(E(X))$, if $Y(X)$ is positive second derivative; and $E(Y(X)) < Y(E(X))$ if $Y(X)$ is negative second derivative. Most packages for predicting responses make distribution assumptions about the predictors, (for example, conditioning at the mean value of the predictor), leading to potential biasness if the particular predictor is not well represented, or as a result of nonlinear averaging. We consider the following approaches for bias correction:

- Population averaging

	- Whole population
	- Quantiles

- Distributional conditioning
- Second-order correction

We implement and apply these methods in the context of both simple generalized linear and mixed effect models, using simulated data sets.

\section{Simple fixed effect model}

\begin{align*}
\mathrm{logit(status} = 1) &= \eta \\ 
\eta &= \beta_0 + \beta_{\mathrm{A}} \mathrm{Age} + \beta_{\mathrm{W}} \mathrm{Wealthindex} \\
\mathrm{Age} &\sim \mathrm{Normal}(0.2, 1) \\
\mathrm{Wealthindex} &\sim \mathrm{Normal}(0, 1) \\
\beta_0 &= 0.7 \\
\beta_{\mathrm{A}} &= 0.2 \\
\beta_{\mathrm{W}} &= 0.5
\end{align*}

```{r sim_simple}
N <- 1e4
beta0 <- 0.7
betaA <- 0.2
betaW <- 0.5

age_max <- 1
age_min <- 0.2
age <- runif(N, age_min, age_max)
# age <- rnorm(N, age_max, age_max)

wealthindex <- rnorm(N, 0, 1)

eta <- beta0 + betaA * age + betaW * wealthindex
sim_df <- (data.frame(age=age, wealthindex=wealthindex, eta=eta)
	%>% mutate(status = rbinom(N, 1, plogis(eta)))
	%>% select(-eta)
)
true_prop <- mean(sim_df$status)
print(true_prop)
head(sim_df)
```

\subsection{Simple logistic model}

```{r simple_logistic}
simple_mod <- glm(status ~ age + wealthindex, data = sim_df, family="binomial")
```

Coefficient plots

```{r simple_coef, fig.cap="Coefficient plot for simple logistic model."}
## True beta
true_beta_df <- data.frame(term=c("Intercept", "age", "wealthindex")
	, estimate=c(beta0, betaA, betaW)
)

## Tidy coef estimates
coef_df <- (broom::tidy(simple_mod, conf.int=TRUE)
#	%>% dotwhisker::by_2sd(sim_df)
	%>% mutate(term = gsub("\\(|\\)", "", term))
)
print(coef_df)

simple_coef_plot <- (plotEsize(coef_df) 
	+ geom_point(data=true_beta_df, aes(x=term, y=estimate, colour="Truth"))
	+ labs(colour="Type")
	+ scale_colour_manual(values=c("black", "blue"))
)
print(simple_coef_plot)
```

Effect sizes on logit scale:

```{r simple_logit}
coef_df_logit <- (coef_df
	%>% select(term, estimate, conf.low, conf.high)
	%>% group_by(term)
	%>% summarise_all(plogis)
)
print(coef_df_logit)
```

\subsection{Variable predictions}

We consider both \textbf{varpred} and population averaging approach; and then introduce bias correction to \textbf{varpred} predictions.

```{r pop_ave_fun, echo=FALSE}
popavefun <- function(mod, focal, non.focal, level=0.95, steps=100, modelname="quantiles", ...) {
	mf <- model.frame(mod)
	mm <- (mf
		%>% select_at(c(focal, non.focal))
	)
	quant <- seq(0, 1, length.out=steps)
	mm <- sapply(c(focal, non.focal), function(x)quantile(mm[,x], quant), simplify = FALSE)
	mm <- do.call("expand.grid", mm)
	vc <- vcov(mod)
	mm2 <- model.matrix(formula(mod)[c(1,3)], mm)
	linpred <- as.vector(mm2 %*% coef(mod))
	pse_var <- sqrt(rowSums(mm2 * t(tcrossprod(data.matrix(vc), mm2))))
	z.val <- qnorm(1 - (1 - level)/2)
	pred_df <- (mm
		%>% select_at(focal)
		%>% mutate(fit = linpred
			, lwr = plogis(fit - z.val*pse_var)
			, upr = plogis(fit + z.val*pse_var)
			, fit = plogis(fit)
		)
		%>% group_by_at(focal)
		%>% summarise_at(c("fit", "lwr", "upr"), mean)
		%>% mutate(model = modelname, se=NA)
	)
	return(pred_df)
}

binfun <- function(mod, focal, non.focal, bins=50, ...) {
	mf <- model.frame(mod)
	mm <- (mf
		%>% select_at(c(focal, non.focal))
	)
	check_df <- (mf
		%>% arrange_at(focal)
		%>% mutate(bin=ceiling(row_number()*bins/nrow(.)))
		%>% group_by(bin)
		%>% summarise_all(mean)
		%>% mutate(model="binned")
	)
	return(check_df)
}

```

- Age

```{r simple_vareffects_age} 
## varpred way
simple_vareff_age <- varpred(simple_mod, "age", isolate=FALSE, modelname="Distribution")

## Pop. average
### Quantiles
simple_vareff_age_quant <- varpred(simple_mod, "age", isolate=FALSE, pop.ave="quantile", modelname="quantiles")
### Population
simple_vareff_age_pop <- varpred(simple_mod, "age", isolate=FALSE, pop.ave="population", modelname="Population")
binned_df <- binfun(simple_mod, "age", "wealthindex", bins=15)

vareff_age <- simple_vareff_age
vareff_age$preds <- do.call("rbind", list(vareff_age$preds, simple_vareff_age_quant$preds, simple_vareff_age_pop$preds))
age_plot <- (plot(vareff_age)
 	+ labs(y="Prob. of improved \n service", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ geom_point(data=binned_df, aes(x=age, y=status, color="binned"))
	+ scale_colour_manual(values=c("grey", "blue", "red", "black"))
	+ theme(legend.position="bottom")
)
```

- Wealth index

```{r simple_vareff_wealthindex, fig.cap = "A comparison of various bias-correction approaches -- Distribution approach is based on averaging of non-focal predictors; Population approach involves using the observed values of non-focal predictors together with the quantiles of the focal predictors; while quantiles approach involves sampling (or picking quantiles of) the observed values of both focal and non-focal predictors. Except for the distribution approach, population and quantiles approaches give very close estimates."}
# Wealth index
## varpred 
simple_vareff_wealthindex <- varpred(simple_mod, "wealthindex", isolate=FALSE, modelname="Distribution")

## Pop. average
### Quantiles
simple_vareff_wealthindex_quant <- varpred(simple_mod, "wealthindex", isolate=FALSE, pop.ave="quantile", modelname="quantiles")
### Population
simple_vareff_wealthindex_pop <- varpred(simple_mod, "wealthindex", isolate=FALSE, pop.ave="population", modelname="Population")
binned_df <- binfun(simple_mod, "wealthindex", "age")

vareff_wealthindex <- simple_vareff_wealthindex
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, simple_vareff_wealthindex_quant$preds, simple_vareff_wealthindex_pop$preds))
wealthindex_plot <- (plot(vareff_wealthindex)
 	+ labs(y="", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ geom_point(data=binned_df, aes(x=wealthindex, y=status, color="binned"))
	+ scale_colour_manual(values=c("grey", "blue", "red", "black"))
	+ theme(legend.position="bottom")
)

ggarrange(age_plot, wealthindex_plot, common.legend=TRUE)
```

The observed population average is `r true_prop` while the estimated population averages (similar estimates with `varpred`) are:

- age: quantiles `r mean(simple_vareff_age_quant$preds$fit)`; population `r mean(simple_vareff_age_pop$preds$fit)`; distribution `r mean(simple_vareff_age$preds$fit)`
- wealthindex: quantiles `r mean(simple_vareff_wealthindex_quant$preds$fit)`; population `r mean(simple_vareff_wealthindex_pop$preds$fit)`; distribution `r mean(simple_vareff_wealthindex$preds$fit)`

\section{Random effect model}

```{r sim_randeff}
# Simulation parameters
nHH <- 100	# Number of HH (primary units) per year

nyrs <- 50	# Number of years to simulate
yrs <- 2000 + c(1:nyrs) # Years to simulate
N <- nyrs * nHH

## HH random effect sd
hhSD <- 0.5

# Generate dataset template
temp_df <- (data.frame(hhid = rep(c(1:nHH), each = nyrs)
		, years = rep(yrs, nHH)
		, age = runif(n=N, age_min, age_max)
		, wealthindex = rnorm(n = N, 0, 1)
	)
)

# Simulate HH-level random effects (residual error)
hhRE <- rnorm(nHH, hhSD)
temp_df$hhRE <- hhRE[temp_df$hhid]

sim_df <- (temp_df
	%>% mutate(eta = beta0 + betaA * age + betaW * wealthindex + hhRE
		, status = rbinom(N, 1, plogis(eta))
	)
	%>% select(-eta)
)
true_prop_reff <- mean(sim_df$status)
print(true_prop)
print(head(sim_df, 10))
```

\subsection{Fit model}

```{r model_reff}
reff_mod <- glmmTMB(status ~ age + wealthindex + (1|hhid)
	, data = sim_df
	, family = binomial(link = "logit")
)
```

```{r reff_coef, fig.cap="Mixed model coefficient estimates."}
## Tidy coef estimates
reff_coef_df <- (broom.mixed::tidy(reff_mod, conf.int=TRUE)
	%>% mutate(term = gsub("\\(|\\)", "", term))
	%>% filter(effect=="fixed")
)
print(reff_coef_df)

reff_coef_plot <- (plotEsize(reff_coef_df) 
	+ geom_point(data=true_beta_df, aes(x=term, y=estimate, colour="Truth"))
	+ labs(colour="Type")
	+ scale_colour_manual(values=c("black", "blue"))
)
print(reff_coef_plot)
```

Variable effect plots

- Age

```{r reff_vareffects_age}
## varpred way
reff_vareff_age <- varpred(reff_mod, "age", isolate=FALSE, modelname="Distribution")

## Pop. average
### Quantiles
reff_vareff_age_quant <- varpred(reff_mod, "age", isolate=FALSE, pop.ave="quantile", include.re=TRUE, modelname = "quantiles")
### Population
reff_vareff_age_pop <- varpred(reff_mod, "age", isolate=FALSE, pop.ave="population", include.re=TRUE, modelname = "Population")
binned_df <- binfun(reff_mod, "age", "wealthindex", bins=15)

vareff_age <- reff_vareff_age
vareff_age$preds <- do.call("rbind", list(vareff_age$preds, reff_vareff_age_quant$preds, reff_vareff_age_pop$preds))
age_plot <- (plot(vareff_age)
 	+ labs(y="Prob. of improved \n service", colour="Model")
	+ geom_hline(yintercept=true_prop_reff, lty=2, colour="grey")
	+ geom_point(data=binned_df, aes(x=age, y=status, color="binned"))
	+ scale_colour_manual(values=c("grey", "blue", "red", "black"))
	+ theme(legend.position="bottom")
)
```

- Wealth index

```{r reff_vareff_wealthindex, fig.cap = "Distribution based approach over-estimates the predictions in age and wealth index."} 
# Wealth index
## varpred
reff_vareff_wealthindex <- varpred(reff_mod, "wealthindex", isolate=FALSE, modelname="Distribution")
binned_df <- binfun(reff_mod, "wealthindex", "age")

## Pop. average
### Quantiles
reff_vareff_wealthindex_quant <- varpred(reff_mod, "wealthindex", isolate=FALSE, pop.ave="quantile", include.re=TRUE, modelname="quantiles")
### Population
reff_vareff_wealthindex_pop <- varpred(reff_mod, "wealthindex", isolate=FALSE, pop.ave="population", include.re=TRUE, modelname="Population")

vareff_wealthindex <- reff_vareff_wealthindex
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, reff_vareff_wealthindex_quant$preds, reff_vareff_wealthindex_pop$preds))
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, reff_vareff_wealthindex_quant$preds, reff_vareff_wealthindex_pop$preds))
wealthindex_plot <- (plot(vareff_wealthindex)
 	+ labs(y="", colour="Model")
	+ geom_hline(yintercept=true_prop_reff, lty=2, colour="grey")
	+ geom_point(data=binned_df, aes(x=wealthindex, y=status, color="binned"))
	+ scale_colour_manual(values=c("grey", "blue", "red", "black"))
	+ theme(legend.position="bottom")
)

ggarrange(age_plot, wealthindex_plot, common.legend=TRUE)
```