### 2021-09-03 aprroximation_bias_correction.rmd
Autocommit

---
title: "Taylor approximation - based methods for bias correction"
author: Bicko, Jonathan & Ben
date: "2021 Aug 11 (Mon)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")

library(glmmTMB)
library(dplyr)
library(ggplot2)
library(vareffects); varefftheme()
library(ggpubr)
set.seed(1011)

prob2logit <- function(x)log(x / (1 - x))
```

\section{Introduction}

We consider two approaches:

\begin{itemize}
\item Second order Taylor approximation
\item Delta method
\end{itemize}

and then compare our approximations to _closed form integral of Gaussian over logistic curve_.

\subsection{Second order Taylor approximation}

We consider Taylor series approximation approach for bias correction. Suppose $f(.)$ represents a back-transformation function, $\mu$ the mean and $\sigma^2_\mu$ on the linear predictor scale (**univariate case** // **what happens in multivariate case?**) predictor. Then using Taylor expansion around $\mu$, the approximation on the original scale is given by

\[
f(\mu) + \frac{1}{2}\sigma^2_{\mu}f''(\mu)
\]

To start with, we consider logistic function

\[
g(\mu) = \frac{1}{1 + \exp(-\mu)},
\]
with the first and second derivatives give by

\[
g'(\mu) = g(\mu)(1 - g(\mu))
\]
and

\[
g''(\mu) = g(\mu)g(-\mu)(1 - 2g(\mu))
\]
respectively.

```{r taylor_fun}
taylorapprox <- function(fun, var="x", mu, sigma, ...) {
	ff <- as.expression(substitute(fun))
	x <- mu
	untrans <- eval(ff, list(x))
	der <- D(ff, var)		# 1s derivative
	der2 <- D(der, var)	# 2nd derivative
	der2 <- eval(der2, list(x))
	corrected <- untrans + der2*sigma^2/2
	return(corrected)
}
```

\subsection{Delta method}

Consider a first order Taylor approximation of $f(.)$ about the mean $\mu$, evaluated at the random variable $x_i$ defined as:

\[
f(x_i) \approx f(\mu) + f'(x_i)(x_i - \mu)
\]
with a variance of

\[
\mbox{Var}\left(f(x_i)\right) = f'(\mu)\mbox{Var}(x_i)f'(\mu)
\]

```{r deltaapprox}
deltaapprox <- function(fun, var="x", z, ...) {
	ff <- as.expression(substitute(fun))
	der <- D(ff, var)
	x <- z
	ymu <- eval(ff, list(x))
	x <- mean(z)
	der <- eval(der, list(x))
	y <- ymu + der * (z - x)
	yvar <- der * var(z) * der		# Check with BB
	out <- list(fx = y, var = yvar)
	return(out)
}
```

Let us consider a simple univariate case

```{r simple_univariate}
b0 <- 6
b1 <- 5
N <- 100
x <- scale(rnorm(N))
eta <- b0 + b1*x

taylor <- taylorapprox(fun=1/(1+exp(-x)), mu=mean(eta), sigma=sd(eta))
delta <- deltaapprox(fun=1/(1+exp(-x)), z=eta)
normal_approx <- logitnorm::momentsLogitnorm(mu= mean(eta), sigma = sd(eta))
out <- c(NULL
	, truth = mean(plogis(eta))
	, naive = plogis(mean(eta))
	, taylor = taylor
	, delta = mean(delta$fx)
	, normal = normal_approx[[1]]
)
print(out)
delta$var
normal_approx[[2]]
```

Simple logistic model example

```{r simple_sim}
N <- 1e4
beta0 <- 0.7
betaA <- 0.2
betaW <- 0.5

age_max <- 1
age_min <- 0.2
age <- runif(N, age_min, age_max)
# age <- rnorm(N, age_max, age_max)

wealthindex <- rnorm(N, 0, 1)

eta <- beta0 + betaA * age + betaW * wealthindex
sim_df <- (data.frame(age=age, wealthindex=wealthindex, eta=eta)
	%>% mutate(status = rbinom(N, 1, plogis(eta)))
	%>% select(-eta)
)
true_prop <- mean(sim_df$status)
print(true_prop)
```

```{r simple_logistic}
simple_mod <- glm(status ~ age + wealthindex, data = sim_df, family="binomial")
```

Applying `taylor`, `delta` and `normal approximation` method

```{r generate_pred}
genpred <- function(mod, focal, non.focal, level=0.95, steps=100, pop=FALSE, bias.adjust=c("none", "delta", "normal", "taylor"), modelname="none", ...) {

	bias.adjust <- match.arg(bias.adjust)
	mf <- model.frame(mod)
	mm <- (mf
		%>% select_at(c(focal, non.focal))
	)
	quant <- seq(0, 1, length.out=steps)
	mm1 <- sapply(focal, function(x)as.vector(quantile(mm[,x], quant)), simplify = FALSE)
	if (pop) {
		mm1[[non.focal]] <- mm[[non.focal]]
		mm <- do.call("expand.grid", mm1)
		mm <- model.matrix(formula(mod)[c(1,3)], mm)
	} else {
		mm2 <- sapply(non.focal, function(x)mean(mm[,x]), simplify = FALSE)
		mm <- do.call("data.frame", c(mm1, mm2))
		mm <- model.matrix(formula(mod)[c(1,3)], mm)
	}
	linpred <- as.vector(mm %*% coef(mod))
	out <- (mm
		%>% data.frame()
		%>% select_at(focal)
		%>% mutate(lp = linpred, model=modelname)
	)

	if (bias.adjust=="delta") {
		out <- (out
			%>% mutate(fit=deltaapprox(fun=1/(1+exp(-x)), z=lp)$fx)
		)
	} else if (bias.adjust=="taylor") {
		vc <- vcov(mod)
		pse_var <- sqrt(rowSums(mm * t(tcrossprod(data.matrix(vc), mm))))
		lp <- out$lp
		fit <- unlist(lapply(1:length(lp), function(i){
			taylorapprox(fun=1/(1+exp(-x)), mu=lp[[i]], sigma=pse_var[[i]])
		}))
		out <- (out
			%>% select(-lp)
			%>% mutate(fit = fit) 
		)
	}	else if (bias.adjust=="normal") {
		vc <- vcov(mod)
		pse_var <- sqrt(rowSums(mm * t(tcrossprod(data.matrix(vc), mm))))
		lp <- out$lp
		fit <- unlist(lapply(1:length(lp), function(i){
			logitnorm::momentsLogitnorm(mu = lp[[i]], sigma = pse_var[[i]])[[1]]
		}))
		out <- (out
			%>% select(-lp)
			%>% mutate(fit = fit) 
		)
	} else if (bias.adjust=="none") {
		out <- (out
			%>% mutate(fit = plogis(lp)) 
		)
	}
	if (pop) {
		out <- (out
			%>% group_by_at(focal)
			%>% summarise_at("fit", mean)
			%>% mutate(model=modelname)
		)
	}
	return(out)
}

### Population averaging
#pop_none <- genpred(simple_mod, "age", "wealthindex", pop=TRUE, bias.adjust="none", modelname="pop-none")
#pop_delta <- genpred(simple_mod, "age", "wealthindex", pop=TRUE, bias.adjust="delta", modelname="pop-delta")
#
### Averaged // centered??
#centered_none <- genpred(simple_mod, "age", "wealthindex", pop=FALSE, bias.adjust="none", modelname="centered-none")
#centered_delta <- genpred(simple_mod, "age", "wealthindex", pop=FALSE, bias.adjust="delta", modelname="centered-delta")
#centered_taylor <- genpred(simple_mod, "age", "wealthindex", pop=FALSE, bias.adjust="taylor", modelname="centered-taylor")
#
```

```{r generate_pred2}
genpred2 <- function(mod, focal, level=0.95, steps=100, modelname="none", ...) {

	mm <- model.matrix(mod)
	quant <- seq(0, 1, length.out=steps)
	xvals <- as.vector(quantile(mm[, focal], quant))
	vnames <- vareffects:::get_vnames(mod)$termnames
	check <- !vnames %in% focal
	non.focal.terms <- vnames[check]
	betas <- coef(mod)
	non.focal.coefs <- betas[check]
	focal.coefs <- betas[!check]
	lp.non.focal <- as.vector(as.matrix(mm[, non.focal.terms, drop=FALSE]) %*% non.focal.coefs)
	lp.all <- unlist(lapply(xvals, function(x){
		mu <- mean(plogis(as.vector(c(x %*% focal.coefs) + lp.non.focal)))
		return(mu)
	}))

	out <- list(xx = xvals, xx2 = lp.all)
	return(out)
}

x1 <- genpred2(simple_mod, "wealthindex")
mean(x1$xx2)

```

```{r simple_preds, echo=FALSE}
#p1 <- (ggplot(pop_none, aes(x=age, y=fit, colour=model))
#	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
#	+ geom_line()
#	+ geom_line(data=pop_delta, aes(x=age, y=fit, colour=model))
#	+ geom_line(data=centered_none, aes(x=age, y=fit, colour=model))
#	+ geom_line(data=centered_delta, aes(x=age, y=fit, colour=model))
#	+ geom_line(data=centered_taylor, aes(x=age, y=fit, colour=model))
#	+ scale_colour_manual(values=c("blue", "red", "black", "orange", "purple"))
#	+ theme(legend.position="right")
#)
# print(p1)
```