### 2021-10-07 compare_emmeans_varpred.rmd
Emmeans vs varpreds

---
title: "Bias correction in GLMs"
author: Bicko, Jonathan & Ben
date: "2021 Oct 06 (Wed)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
bibliography: bias_correction.bib
link-citations: yes
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "", message=FALSE, warning=FALSE, echo=FALSE)

library(shellpipes)
library(glmmTMB)
library(dplyr)
library(ggplot2)
library(vareffects); varefftheme()
library(emmeans)

commandEnvironments()
makeGraphics()

set.seed(2121)

vplots <- vareffects:::plot.vareffects

```

\section{Continuous predictors}

\subsection{No interaction}

```{r cont_no_interaction}
quants <- seq(0, 1, length.out=100)
focal_levels <- quantile(sim_df_cni$age, quants)
truth = c(truth=mean(sim_df_cni$hhsize))
print(truth)
compare_cni_plots <- (combinepreds(mod_cni, focal="age"
		, funs=c("emmeans", "varpred")
		, at=list(age=focal_levels)
		, x.var="age"
		, plotit=TRUE
	)
	+ geom_vline(aes(xintercept=mean(focal_levels)), lty=2, colour="grey")
)
print(compare_cni_plots)
```

\subsection{Interaction between non-focal predictors}

```{r cont_nwi_nf_interaction}
quants <- seq(0, 1, length.out=100)
focal_levels <- quantile(sim_df_wi_nf$age, quants)
truth = c(truth=mean(sim_df_wi_nf$hhsize))
print(truth)
compare_wi_nf_plots <- (combinepreds(mod_wi_nf, focal="age"
		, funs=c("emmeans", "varpred")
		, at=list(age=focal_levels)
		, x.var="age"
		, plotit=TRUE
	)
	+ geom_vline(aes(xintercept=mean(focal_levels)), lty=2, colour="grey")
)
print(compare_wi_nf_plots)
```