### 2021-10-28 compare_emmeans_varpred.rmd
Categorical interactions

---
title: "emmeans and varpred comparison"
author: Bicko, Jonathan & Ben
date: "2021 Oct 06 (Wed)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
bibliography: bias_correction.bib
link-citations: yes
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "", message=FALSE, warning=FALSE, echo=FALSE)

library(shellpipes)
library(glmmTMB)
library(dplyr)
library(ggplot2)
library(vareffects); varefftheme()
library(emmeans)
library(effects)

commandEnvironments()
makeGraphics()

set.seed(2121)

within.category <- TRUE

```

\section{Continuous predictors}

\subsection{No interaction}

```{r cont_no_interaction, fig.cap="Continuous: no interaction"}
quants <- seq(0, 1, length.out=100)
focal_levels <- quantile(sim_df_cni$age, quants)
print("Truth:")
print(true_prop_cni)
compare_cni_plots <- (combinepreds(mod_cni, focal="age"
		, funs=c("emmeans", "varpred")
		, at=list(age=focal_levels)
		, x.var="age"
		, plotit=TRUE
	)
	+ geom_vline(aes(xintercept=mean(focal_levels)), lty=2, colour="grey")
)
print(compare_cni_plots)
```

\subsection{Interaction between non-focal predictors}

```{r cont_wi_nf_interaction, fig.cap="Continuous: non-focal predictors interaction"}
focal_levels <- quantile(sim_df_wi_nf$age, quants)
print("Truth:")
print(true_prop_wi_nf)
compare_wi_nf_plots <- (combinepreds(mod_wi_nf, focal="age"
		, funs=c("emmeans", "varpred")
		, at=list(age=focal_levels)
		, x.var="age"
		, plotit=TRUE
	)
	+ geom_vline(aes(xintercept=mean(focal_levels)), lty=2, colour="grey")
)
print(compare_wi_nf_plots)
```

\subsection{Interaction between focal and non-focal predictors}

```{r cont_wi_f_interaction, fig.cap="Continuous: interaction betweem focal and non-focal predictors"}
focal_levels <- quantile(sim_df_wi_f$age, quants)
print("Truth:")
print(true_prop_wi_f)
compare_wi_f_plots <- (combinepreds(mod_wi_f, focal="age"
		, funs=c("emmeans", "varpred")
		, at=list(age=focal_levels)
		, x.var="age"
		, nesting="wealthindex %in% age"
		, plotit=TRUE
	)
	+ geom_vline(aes(xintercept=mean(focal_levels)), lty=2, colour="grey")
)
print(compare_wi_f_plots)
```

\section{Categorical predictors}

\subsection{No interaction}

```{r catni, fig.cap="Categorical: no interaction"}
compare_catni_plots <- (combinepreds(mod_catni, focal="gender"
		, funs=c("emmeans", "varpred")
		, x.var="gender"
		, within.category=within.category
		, x.var.factor=TRUE
		, plotit=TRUE
	)
	+ geom_point(data=focal_prop_catni, aes(x=gender, y=hhsize), colour="black")
)
print(compare_catni_plots)
```

\section{Why the differences?}

We actually don't know which method is correct. Let us consider a simple simulation:

- Outcome: hhsize
- Predictors:
	- Gender (40\% Males)
	- Age: Females are slightly older

```{r sim_compare, echo=TRUE}
## Simulation
N <- 100
extraAge <- 0.4
meanAge <- 0.5
sdAge <- 1
prop <- c(0.6, 0.4)
betas <- c(1.5, 0.5, 0.5)
gender <- sample(c("Female", "Male"), N, replace = TRUE, prob = prop)
df <- (data.frame(gender=gender)
	%>% mutate(muAge=ifelse(gender=="Female", meanAge+extraAge, meanAge)
		, age=rnorm(N, muAge, sdAge)
	)
	%>% select(-muAge)
)

mm <- model.matrix(~gender+age, df)
df$hhsize <- rnorm(N, mean=as.vector(mm %*% betas), sd=1)

## Model
mod <- lm(hhsize~gender+age, data=df)
print(mod)
```

Let us take a look at the observed marginals:

```{r observed_margins, echo=TRUE}
observed_margins <- (df
	%>% group_by(gender)
	%>% summarise_all(mean)
)
observed_margins
```

- **varpred** constructs model matrix by averaging _age_ within the levels of _gender_. The population average is the weighted (by the observed proportions) average of these averages:

```{r varpred_mm, echo=TRUE}
varpred_pred <- varpred(mod, "gender", within.category=TRUE, returnall=TRUE)
varpred_mm <- varpred_pred$raw$model.matrix
print(varpred_mm)
```

Estimates:
```{r varpred_pred}
print(varpred_pred)
```

By hand calculation:

```{r varpred_hand, echo=TRUE}
varpred_mm %*% coef(mod)
```

- **emmeans** uses the population average and seems not apply the weights:

```{r emmeans_mm, echo=TRUE}
emmeans_grid <- ref_grid(mod)
emmeans_mm <- emmeans_grid@grid
print(emmeans_mm)
```

Estimates:
```{r emmeans_pred}
emmeans_pred <- emmeans(emmeans_grid, spec=~gender)
print(emmeans_pred)
```

By hand calculation:

```{r emmeans_hand, echo=TRUE}
emmeans_mm <- emmeans_pred@linfct
print(emmeans_mm)

as.matrix(emmeans_mm) %*% coef(mod)
```

Mathematically, in **varpred**

\begin{align}
\mathbb{E}(Y|X) =
\begin{cases}
(\beta_0 + \alpha) + \beta_1 \bar{{Age_{M}}},& \text{ if } gender = Male\\\\
\beta_0 + \beta_1 \bar{{Age_{F}}},&  \text{ if } gender = Female\
\end{cases}
\end{align}

**emmeans**

\begin{align}
\mathbb{E}(Y|X) =
\begin{cases}
(\beta_0 + \alpha) + \beta_1 \bar{{Age}},& \text{ if } gender = Male\\\\
\beta_0 + \beta_1 \bar{{Age}},&  \text{ if } gender = Female\
\end{cases}
\end{align}

\section{Dealing with categorical iterations}

```{r categorical_toy, echo=TRUE}
df <- (expand.grid(services=c("water", "garbage", "toilet")
		, gender=c("Male", "Female")
		, statusP=c("Base year", "Not observed", "Unimproved", "Improved")
	)
	%>% mutate(age = rnorm(n())
		, status=rbinom(n(), 1, plogis(rnorm(n(), 3, 5)))
	)
)
head(df)
```

Model:

```{r toy_model, echo=TRUE}
toy_mod <- glm(status~-1+services+(age+gender+statusP):services
	, data=df
	, family="binomial"
	, contrasts=list(services="contr.sum")
)
```

\subsection{varpred}

- Averaging all the columns of model matrix?

```{r varpred_all_av}
varpred_all_av <- varpred(toy_mod, "services", handle.inter="emmeans", returnall=TRUE)
print(varpred_all_av$raw$model.matrix)
```

- On baseline level?

```{r varpred_no_av}
varpred_no_av <- varpred(toy_mod, "services", handle.inter="effects", returnall=TRUE)
print(varpred_no_av$raw$model.matrix)
```

\subsection{emmeans}

```{r emmeans_toy}
emmeans_toy <- emmeans(toy_mod, specs=~services, nesting=NULL, weighting="equal")
print(emmeans_toy@linfct)
```

\subsection{Effects}

```{r effects_toy}
effects_toy <- Effect("services", toy_mod)
print(effects_toy$model.matrix)
```