### 2021-11-02 short_lab_report.rmd
Autocommit

---
title: "Predictor effects and Bias correction in GL(M)M"
author: Bicko, Ben & Jonathan
date: "2021 Oct 29 (Fri)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
bibliography: bias_correction.bib
link-citations: yes
---

```{r, echo=FALSE}
library(ggplot2)

library(vareffects); varefftheme()

set.seed(9991)
load("variable_predictions_objs.rda")
load("binom_correlated_preds.rda")
```

We want:

\[
E(y|\{X_f, \bar{X}_{\{n\}}\})
\]

Challenges:

\begin{enumerate}
\item choice of representative values of \emph{focal} predictor(s) and appropriate ``model center" for \emph{non-focal} predictors especially in multivariate models
\item  propagation of uncertainty ([how] can we incorporate uncertainty in nonlinear components of (G)LMMs? Should we exclude variation in non-focal parameters?)
\item bias in the expected mean prediction induced by the nonlinear transformation of the response variable (especially in GL(M)M)
\end{enumerate}

```{r echo=FALSE, fig.height=10}
print(pred_age_cont_plots)
```

```{r echo=FALSE, fig.height=10}
print(pred_bin_corr_plots)
```