### 2021-11-09 bias_correction_methods.rmd
A lot of pieces

---
title: "Bias correction in GLMs"
author: Bicko, Jonathan & Ben
date: "2021 Jun 21 (Mon)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "", message=FALSE, warning=FALSE)

library(shellpipes)
library(dplyr)
library(ggplot2)
library(vareffects); varefftheme()

commandEnvironments()
makeGraphics()

set.seed(9991)
```

\newcommand{\nset}[1]{#1_{\{n\}}}
\newcommand{\yref}{y_{\textrm{ref}}}
\newcommand{\cdist}{{D(\nset{x}|x_f)}}
\newcommand{\cdistprime}{{D(\nset{x}|x_{f'})}}
\newcommand{\xfprime}{x_{f'}}

\newcommand{\bX}{{\mathbf X}}
\newcommand{\bbeta}{{\boldsymbol \beta}}
\newcommand{\boldeta}{{\boldsymbol \eta}}

Notation :

- $x_f$: a value of the focal predictor
- $\nset{x}$: a vector of values of the non-focal predictors for a particular observation
- $\eta(x_f, \nset{x}) = \beta_f x_f + \sum \nset{\beta} \nset{x}$ = linear predictor (e.g. prediction on the log-odds scale)
- $g^{-1}()$: inverse-link function (e.g. logistic)
- $\cdist$: distribution of the non-focal predictors conditional on a particular value of the focal predictor
- $\beta_{fi}$: the coefficient describing the interaction(s) of the focal and non-focal parameters

_\

\section{Population-based approach for bias correction}

\subsection{Binned non-focal linear predictor}

To implement this:
\begin{itemize}
\item compute linear predictor of the non-focal predictors, $\nset{\eta} = \sum \nset{\beta} \nset{x}$
\item find a list of vectors of observations of $\nset{\eta}$ associated with each value (bin) of the focal predictor, $\nset{{\eta_j}}$, $j = 1, 2, \cdots$
\item for each $\nset{{\eta_j}}$:
\begin{itemize}
\item compute $\hat{y}_j  = \textrm{mean} ~ g^{-1} \left(\beta_f x_{j_f} + \nset{{\eta_j}}\right)$
\end{itemize}
\end{itemize}

If we compute the individual back-transformed predictions for a poorly sampled/finely spaced set of focal values, we will get a noisy prediction line as the values of the non-focal predictors shift across the focal values. Simple example: suppose everyone below the median age has wealth index $w_1$, everyone above the median has $w_2$.  Then the predicted value will have a discontinuity at the median age. We can deal with this by taking bigger bins (a form of smoothing), or by post-smoothing the results (by loess, for example). The principled form of this would be to assume/recognize that our uneven distribution of observed non-focal predictors actually represents a sample of a distribution that will vary \emph{smoothly} as a function of the focal predictor.

\subsection{Whole population non-focal linear predictor}

Instead of binning non-focal associated with particular level of focal predictor, we add the overall contribution of the non-focal predictor to the corresponding value of the focal linear predictor. In particular:

\begin{itemize}
\item compute linear predictor of the non-focal predictors, $\nset{\eta} = \sum \nset{\beta} \nset{x}$
\item for every value of the focal predictor, $x_{j_f}$:
\begin{itemize}
\item compute $\hat{y}_j  = \textrm{mean} ~ g^{-1} \left(\beta_f x_{j_f} + \nset{\eta}\right)$
\end{itemize}
\end{itemize}

\section{Delta method}

Suppose that we have a response $Y$ and the transformed response on the link scale, $\eta$ and an inverse-link function $h(.) = g^{-1}(.)$. We consider $\eta$ as a one-dimensional random variable with mean ($\mu_{\eta}$) and standard deviation ($\sigma_{\eta}$). We are interested in the expected value of $h(\eta)$, i.e., the back-transformed values so that $Y = h(\eta)$. More specifically, based on a second-order Taylor expansion

\begin{align*}
Y \approx h(\mu_{\eta}) + h'(\eta)(\eta - \mu_{\eta}) + \frac{1}{2}h''(\eta)(\eta - \mu_{\eta})^2
\end{align*}

so that

\begin{align*}
E(Y) = h(\eta) + \frac{1}{2}h''(\mu_{\eta})\sigma^2_{\eta}
\end{align*}

\section{McCulloch and Searle (2001)}
\[
E(Y) \approx h\left(\eta-\frac{0.5\sigma^2\tanh(\eta(1 + 2\exp(-0.5\sigma^2)))}{6}\right)
\]

\section{Diggle et al. (2004)}
\[
E(Y) \approx h\left(\frac{\eta}{\sqrt{1 + \left(\frac{16\sqrt{3}}{15\pi}\right)^2\sigma^2}}\right)
\]