### 2021-11-09 understanding_mm_special.rmd

---
title: "How to we handle the categorical predictors?"
author: Bicko, Jonathan & Ben
date: "2021 Nov 06 (Sat)"
compact-title: false
output:
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.pos = "H"
	, out.extra = ""
	, message=FALSE
	, warning=FALSE
	, echo=FALSE
	, comment=""
)

library(shellpipes)
library(dplyr)
library(ggplot2)
library(vareffects); varefftheme()
library(emmeans)
library(effects)

commandEnvironments()
makeGraphics()

set.seed(2121)

load("variable_predictions_funs.rda")

cols <- c(Effect="black", varpred="red", emmeans="blue")
```

\section{Two categorical, single continuous predictors}

\begin{align}\label{sim:lm_no_interaction_cat}
\mathrm{hh~size} &= \beta_0 + \beta_{\mathrm{A}}\mathrm{Age}_i + \beta_{[\mathrm{r}]}\mathrm{Religion_i} + \beta_{[\mathrm{g}]}{\mathrm{Gender}_i} + \beta_{[\mathrm{rg}]}\mathrm{Gender_i
\times Religion_i} + \epsilon_i \nonumber\\
\mathrm{Age}_i &\sim \mathrm{Normal}(0, 3) \nonumber\\
\mathrm{Prob(Gender_{i})} &= \begin{cases}
   0.4 & \mathrm{i=Female} \nonumber\\
   0.6 & \mathrm{i=Male} \nonumber\\
  \end{cases} \nonumber\\
\mathrm{Prob(Religion_{i})} &= \begin{cases}
   0.6 & \mathrm{i=Christian} \nonumber\\
   0.3 & \mathrm{i=Muslim} \nonumber\\
   0.1 & \mathrm{i=Jew} \nonumber\\
  \end{cases} \nonumber\\
\epsilon_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\beta_0 &= 1.5 \nonumber\\
\beta_{\mathrm{A}} &= 0.8 \nonumber\\
\vdots \nonumber \\
i &= 1,\cdots, 100
\end{align}

```{r depend_eff, echo=TRUE}
N <- 100
quant <- seq(0,1,length.out=50)
## Betas
b0 <- 10
b_age <- 0.8
b_jew <- 1
b_muslim <- 0.4
b_male <- 2
b_male_jew <- 0.7
b_male_muslim <- 0.2
betas <- c(b0, b_age, b_jew, b_muslim, b_male, b_male_jew, b_male_muslim)

## Proportions
rel_prob <- c(Christian=0.6, Muslim=0.3, Jew=0.1)
gender_prob <- c(Male=0.6, Female=0.4)

dat <- tibble(age=rnorm(N, 0, 3)
	, rel=sample(c("Christian", "Muslim", "Jew"), size=N, prob=rel_prob, replace=TRUE)
	, gender=sample(c("Male", "Female"), size=N, prob=gender_prob, replace=TRUE)
)
mm <- model.matrix(~1+age+rel*gender, dat)
eta <- as.vector(mm %*% betas)

df <- (dat
	%>% mutate(hhsize=rnorm(N, mean=eta, sd=1))
)
age_at <- quantile(df$age, quant)
head(df)
```

The model:

```{r depend_eff_mod, echo=TRUE}
depend_eff_mod <- lm(hhsize~rel*gender+age, df)
```

\newpage

\subsubsection{gender$\star$religion}

- All the three methods seems to be doing the same thing:
	- product of interacting terms

```{r depend_eff_pred}
depend_eff_pred <- combinepreds(depend_eff_mod, c("emmeans", "varpred", "Effect"), c("gender","rel")
	, x.var = "gender"
	, weights="proportional"
	, x.var.factor = TRUE, plotit = TRUE
)
print(depend_eff_pred 
	+ facet_wrap(~rel, scales = "free_y")
	+ scale_colour_manual(values=cols)
)
```

\newpage

\subsubsection{Main effect of non-interacting predictor}

- If we ignore the interactions:
	- **varpred** and **emmeans** averages the columns of the interaction terms
	- **effects** computes the product of the interacting terms
	- The predictions are similar but slightly different

```{r depend_eff_age}
depend_eff_age <- combinepreds(depend_eff_mod, c("emmeans", "varpred", "Effect"), "age"
	, x.var = "age"
	, at=list(age=age_at)
	, xlevel=list(age=age_at)
	, nesting=NULL
	, weights="proportional"
	, plotit = TRUE
)
cat("Truth: \n", mean(df$hhsize))
print(depend_eff_age
	+ scale_colour_manual(values=cols)
)
```

\section{Unbalanced data}

```{r unbalanced_sim, echo=TRUE}
N <- 100
quant <- seq(0,1,length.out=50)
## Betas
b0 <- 10
b_age <- 0.8
b_jew <- 1
b_muslim <- 0.4
b_male <- 2
b_male_jew <- 0.7
b_male_muslim <- 0.2
betas <- c(b0, b_age, b_jew, b_muslim, b_male, b_male_jew, b_male_muslim)

## Proportions
rel_prob <- c(Christian=0.6, Muslim=0.3, Jew=0.1)
gender_prob <- list(Christian=c(0.97, 0.03)
	, Muslim=c(0.01, 0.99)
	, Jew=c(0.5, 0.5)
)

samplefun <- function(prob, n) {
	out <- sample(c("Male", "Female"), size=n, prob=prob, replace=TRUE)
}

dat <- tibble(age=rnorm(N, 0, 3)
	, rel=sample(c("Christian", "Muslim", "Jew"), size=N, prob=rel_prob, replace=TRUE)
)
dat <- (dat
	%>% group_by(rel)
	%>% mutate(gender=ifelse(rel=="Christian"
			, samplefun(gender_prob[["Christian"]], n())
				, ifelse(rel=="Muslim", samplefun(gender_prob[["Muslim"]], n())
				, samplefun(gender_prob[["Jew"]], n())
			)
		)
	)
	%>% ungroup()
)
head(dat)
mm <- model.matrix(~1+age+rel*gender, dat)
eta <- as.vector(mm %*% betas)

unbalanced_df <- (dat
	%>% mutate(hhsize=rnorm(N, mean=eta, sd=1))
)
age_at <- quantile(unbalanced_df$age, quant)
head(unbalanced_df)

```

The model:

```{r unbalanced_mod, echo=TRUE}
unbalanced_mod <- lm(hhsize~rel*gender+age, unbalanced_df)
```

\newpage

\subsubsection{gender$\star$religion}

- All the three methods seems to be doing the same thing:
	- product of interacting terms

```{r unbalanced_pred}
unbalanced_pred <- combinepreds(unbalanced_mod, c("emmeans", "varpred", "Effect"), c("gender","rel")
	, x.var = "gender"
	, weights="proportional"
	, x.var.factor = TRUE, plotit = TRUE
)
print(unbalanced_pred 
	+ facet_wrap(~rel, scales = "free_y")
	+ scale_colour_manual(values=cols)
)
table(unbalanced_df$rel, unbalanced_df$gender)
```

\newpage

\subsubsection{Main effect of non-interacting predictor}

- If we ignore the interactions:
	- **varpred** and **emmeans** averages the columns of the interaction terms
	- **effects** computes the product of the interacting terms
	- The predictions are similar but slightly different

- CIs excluded for clarity

```{r unbalanced_age}
unbalanced_age <- combinepreds(unbalanced_mod, c("emmeans", "varpred", "Effect"), "age"
	, x.var = "age"
	, at=list(age=age_at)
	, xlevel=list(age=age_at)
	, nesting=NULL
	, weights="proportional"
	, plotit = TRUE
	, ci=FALSE
)
cat("Truth: \n", mean(unbalanced_df$hhsize))
print(unbalanced_age
	+ scale_colour_manual(values=cols)
)
```