### 2022-03-29 manuscript/algebra.Rmd
Autocommit

---
title: Is anchoring better?
---

\newcommand{\yH}{\hat{y}_0}
\newcommand{\yHStarl}{\hat{y}_{0l}^\star}
\newcommand{\yHStarr}{\hat{y}_{0r}^\star}
\newcommand{\xHS}{\bar{x}^\star}
\newcommand{\bH}{\hat{\beta}_1}
\newcommand{\bh}{\hat{\beta}_0}
\newcommand{\sxxb}{\sum{(x_i - \bar{x})^2}}
\newcommand{\sxxbsl}{\sum{(x_i - \xHS_l)^2}}
\newcommand{\sxxbsr}{\sum{(x_i - \xHS_r)^2}}
\newcommand{\var}[1]{\mathrm{var(#1)}}
\newcommand{\ss}{\sigma^2}
\newcommand{\varbH}{\frac{\ss}{\sxxb}}
\newcommand{\varbh}{\ss\left(\frac{1}{n} + \frac{\bar{x}^2}{\sxxb} \right)}

Maybe we can start with simple linear model:

\[
y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

so that

\[
\hat{\beta_0} = \bar{y} - \bH\bar{x} 
\]
with a variance of 
\[
\varbh
\]

and

\[
\bH = \frac{\sum{(x_i - \bar{x})}(y_i - \bar{y})}{\sxxb}
\]
with a variance of 
\[
\varbH
\]
where $\ss$ is the residual variance.

## Prediction

We want to describe the uncertainties associated with the new data point, $x_0$, based on the fitted model, i.e.,

\begin{align}
\var{\yH} &= \var{\bh + \bH x_0}\\
&= \varbh + \varbH - 2\ss\frac{x_0\bar{x}}{\sxxb}\\
&= \ss \left(\frac{1}{n} + \frac{(x_0 - \bar{x})^2}{\sxxb}\right).
\end{align}

The default anchor for prediction is the $\bar{x}$. However, we want to investigate what happens to the $\var{\yH}$ for various choices of anchor, $\xHS_l < \bar{x}$ and $\xHS_r > \bar{x}$. In particular, we want to compare

\begin{align}
\var{\yHStarl} = \ss \left(\frac{1}{n} + \frac{(x_0 - \xHS_l)^2}{\sxxbsl}\right)
\end{align}
and 
\begin{align}
\var{\yHStarr} = \ss \left(\frac{1}{n} + \frac{(x_0 - \xHS_r)^2}{\sxxbsr}\right)
\end{align}
with $\var{\yH}$.